Covariate analysis of ROSMAP reprocessed transcript counts (with TMM normalisation and covariates adjusted)

### 2017-09-15 code/Rmd/MSSM_REPROCESSED_ISOFORM.Rmd
Minor modifications

# Ribosomal bases
p[[4]] = ggplot(COVARIATES, aes(x = BrainRegion.Diagnosis, y = as.numeric(PCT_RIBOSOMAL_BASES))) + geom_boxplot()
p[[4]] = p[[4]] + ggtitle('Fraction of Ribosomal Bases') + my.theme

multiplot(plotlist = p, cols = 2)
```
### Filter genes
Remove genes that have less than 5 tpm counts in at least 50% of samples per BrainRegion x Diagnosis
```{r cpmnormalisation}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('SampleID') %>%
  dlply(.(BrainRegion.Diagnosis), .fun = function(mtd, count){
    processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID],
                                                     MIN_GENE_CPM=5, 
                                                     MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
  processed.counts$filteredExprMatrix$genes
}, COUNT)

>>>>

multiplot(plotlist = p, cols = 3)
```
### Filter genes
Remove genes that have less than 5 tpm counts in at least 50% of samples per BrainRegion x Diagnosis
```{r cpmnormalisation}
genesToAnalyze = COVARIATES %>%
  rownameToFirstColumn('SampleID') %>%
  dlply(.(BrainRegion.Diagnosis), .fun = function(mtd, count){
    processed.counts = getGeneFilteredGeneExprMatrix(count[,mtd$SampleID],
                                                     MIN_GENE_CPM=5, 
                                                     MIN_SAMPLE_PERCENT_WITH_MIN_GENE_CPM=0.5)
  processed.counts$filteredExprMatrix$genes
}, COUNT)

### 2017-09-15 code/Rmd/ROSMAP_REPROCESSED_ISOFORM.Rmd
Minor modifications

---
title: "Covariate analysis of ROSMAP reprocessed isfoform counts (with CQN normalisation and covariates adjusted)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

@@

# Code
CODE <- Folder(name = "ROSMAP Reprocessed Isoform (Normalised)", parentId = parentId)
annotations(CODE) = all.annotations
CODE <- synStore(CODE)
old.acl = synGetEntityACL(CODE@properties$id)
synChangeEntityACL(CODE@properties$id, new.racl)

@@

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('ROSMAP_DLPFC_Counts.tsv', name = 'TPM (raw)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store filtered counts
NEW.COUNTS %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_eCounts.tsv', sep = '\t', row.names=F, quote=F)
ECOUNT_OBJ = File('ROSMAP_DLPFC_eCounts.tsv', name = 'TPM (estimated)', parentId = CODE$properties$id)
annotations(ECOUNT_OBJ) = annotations(CODE)
annotations(ECOUNT_OBJ)$dataSubType = 'estimatedCounts'
ECOUNT_OBJ = synStore(ECOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store logCPM
VOOM.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('ensembl_gene_id') %>%
  write.table(file = 'ROSMAP_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('ROSMAP_DLPFC_logCPM.tsv', name = 'TPM (filtered)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'filteredLTPM'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store design matrix
MODEL1 %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'ROSMAP_DLPFC_Design.tsv', sep = '\t', row.names=F, quote=F)
DM_OBJ = File('ROSMAP_DLPFC_Design.tsv', name = 'Design Matrix', parentId = CODE$properties$id)
annotations(DM_OBJ) = annotations(CODE)
annotations(DM_OBJ)$dataSubType = 'designMatrix'
DM_OBJ = synStore(DM_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
synChangeEntityACL(DM_OBJ@properties$id, new.racl)

@@

# Delete folder acl
synChangeEntityACL(CODE@properties$id, old.acl@resourceAccess)

stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates (90+ ages)                      |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (estimated)                         |  `r ECOUNT_OBJ$properties$id`    |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |
|  Linear Model Fits                          |  `r FIT_OBJ$properties$id`       |

>>>>

---
title: "Covariate analysis of ROSMAP reprocessed transcript counts (with TMM normalisation and covariates adjusted)"
author: "Thanneer Perumal"
date: "`r date()`"
output: html_document
---
```{r knit2synapse, eval=FALSE}
library(synapseClient)
library(knit2synapse) # get the package from devtools::install_github('Sage-Bionetworks/knit2synapse')

@@

# Code
CODE <- Folder(name = "ROSMAP Reprocessed Isoform (Normalised)", parentId = parentId)
annotations(CODE) = all.annotations
CODE <- synStore(CODE)
# old.acl = synGetEntityACL(CODE@properties$id)
synChangeEntityACL(CODE@properties$id, new.racl)

@@

# Store filtered counts
PROCESSED_COUNTS$filteredExprMatrix$counts %>%
  rownameToFirstColumn('Tx.ID') %>%
  write.table(file = 'ROSMAP_DLPFC_Counts.tsv', sep = '\t', row.names=F, quote=F)
COUNT_OBJ = File('ROSMAP_DLPFC_Counts.tsv', name = 'TPM (raw)', parentId = CODE$properties$id)
annotations(COUNT_OBJ) = annotations(CODE)
annotations(COUNT_OBJ)$dataSubType = 'filteredCounts'
COUNT_OBJ = synStore(COUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store filtered counts
NEW.COUNTS %>%
  rownameToFirstColumn('Tx.ID') %>%
  write.table(file = 'ROSMAP_DLPFC_eCounts.tsv', sep = '\t', row.names=F, quote=F)
ECOUNT_OBJ = File('ROSMAP_DLPFC_eCounts.tsv', name = 'TPM (estimated)', parentId = CODE$properties$id)
annotations(ECOUNT_OBJ) = annotations(CODE)
annotations(ECOUNT_OBJ)$dataSubType = 'estimatedCounts'
ECOUNT_OBJ = synStore(ECOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store logCPM
VOOM.GENE_EXPRESSION$E %>%
  rownameToFirstColumn('Tx.ID') %>%
  write.table(file = 'ROSMAP_DLPFC_logCPM.tsv', sep = '\t', row.names=F, quote=F)
LCOUNT_OBJ = File('ROSMAP_DLPFC_logCPM.tsv', name = 'TPM (filtered)', parentId = CODE$properties$id)
annotations(LCOUNT_OBJ) = annotations(CODE)
annotations(LCOUNT_OBJ)$dataSubType = 'filteredLTPM'
LCOUNT_OBJ = synStore(LCOUNT_OBJ, activity = synGetActivity(COV_OBJ$properties$id))

# Store design matrix
DM1$design %>%
  rownameToFirstColumn('SampleID') %>%
  write.table(file = 'ROSMAP_DLPFC_Design.tsv', sep = '\t', row.names=F, quote=F)
DM_OBJ = File('ROSMAP_DLPFC_Design.tsv', name = 'Design Matrix', parentId = CODE$properties$id)
annotations(DM_OBJ) = annotations(CODE)
annotations(DM_OBJ)$dataSubType = 'designMatrix'
DM_OBJ = synStore(DM_OBJ, activity = synGetActivity(COV_OBJ$properties$id))
synChangeEntityACL(DM_OBJ@properties$id, new.racl)

@@

# Delete folder acl
synDeleteEntityACL(CODE@properties$id)

# stopCluster(cl)
```
|  *Results*                                  |  *SynapseID*                     |
|  -----------------------------------------  |   ---------                      |
|  Covariates (90+ ages)                      |  `r COV_OBJ$properties$id`       |
|  Counts (raw)                               |  `r COUNT_OBJ$properties$id`     |
|  Counts (estimated)                         |  `r ECOUNT_OBJ$properties$id`    |
|  Counts (lcpm)                              |  `r LCOUNT_OBJ$properties$id`    |
|  Design Matrix                              |  `r DM_OBJ$properties$id`        |
|  Differential Expression                    |  `r DEXP_OBJ$properties$id`      |
|  Linear Model Fits                          |  `r FIT_OBJ$properties$id`       |
